---
title: "p8105_hw2_ns3923"
Name: Nichcha Subdee (ns3923)
output: github_document
---

**Setup necessary library**
```{r setup, message=FALSE}
library(tidyverse)
```

## Problem 1
Aims: Merging pols-month.csv, unemployment.csv, and snp.csv into a dataframe by using year/month as keys variable

**Import data**
```{r import data, message=FALSE}
polsmonth_df = read_csv(file = "./fivethirtyeight_datasets/pols-month.csv")
unemployment_df = read.csv(file = "./fivethirtyeight_datasets/unemployment.csv")
snp_df = read.csv(file = "./fivethirtyeight_datasets/snp.csv")
```

**Step 1**: Cleaning pols-month.csv

1. Break up `mon` (e.g., 1/15/1947) into integer variables `year`,`month`, and `day`

2. Replace month number with month name

3. Create `president`: taking values `gop` and `dem`

4. Remove `prez_dem` and `prez_gop`, and the day variable
```{r}
polsmonth_df = polsmonth_df |>
    janitor::clean_names() |>
    separate(mon, into = c("year","month","day"), convert = TRUE) |>
    mutate(
           president = ifelse(prez_gop == 1, "gop", "dem"),   
           month = factor(month, levels = 1:12, labels = month.abb)    #Month name
           ) |>
    select(year, month, president,
           starts_with("gov"),
           starts_with("sen"),
           starts_with("rep")) |>
    arrange(year, month)
#Result
polsmonth_df
```




